---
title: "Chap3_Exercise1_BasicDataManiupulation"
author: "Mary Wojtas"
date: "November 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##(0) Load the data into R Studio
```{r}
library(readxl)
refine_original <- read_excel("~/Documents/RProjects/Chapter3Exercise/refine_original.xlsx")
refine <- refine_original
```

##(1) Clean up brand names: transform values in company to lowercase
```{r}
refine$company <- tolower(refine$company)
```

###(1b) Clean up the misspellings
```{r}
refine$company[10] <- "akzo"
refine$company[11] <- "akzo"
refine$company[15] <- "phillips"
refine$company[16] <- "phillips"
refine$company[3:5] <- "phillips"
refine$company[22] <- "unilever"
```
##(2) Create a new column called product_code from Product code/number
## This will separate product code from number
```{r}
library(dplyr)
refine <- mutate(refine, product_code = refine$`Product code / number`[1:25])
```

### (2b) Moves product_code to 1st column
```{r}
refine <- select(refine, product_code, everything())
refine <- select(refine, company, everything())
```

### (2c) Search on p- and substitute just the p for all values in product_code
```{r}
library(stringr)
refine$product_code[1:25] = str_replace(refine$product_code[1:25], "^[:alpha:][:punct:]", str_sub(refine$product_code[1:25], end =1))
```

### (2d) Rename 'Product code/ number' to just product_number
```{r}
refine <- rename(refine, product_number='Product code / number')
```

### (2e) Remove the p- from the values for all columns in product number
```{r}
refine$product_number[1:25] <- str_remove(refine$product_number[1:25], "^[:alpha:][:punct:]")
```

## (3) Add product categories
### (3b) Add a column called product_category: populates new col w/contents from existing column
```{r}
refine <- mutate(refine, product_category = refine$product_code[1:25])
```
### (3c) Populate new column with new value depending on what the pre-existing column had.
```{r}
refine$product_category[1:25] <- case_when(refine$product_category=="p" ~ "SmartPhone", refine$product_category=="x" ~ "Laptop", refine$product_category=="v" ~ "TV", refine$product_category=="q" ~ "Tablet")
```
### (3d) Reorder the columns
```{r}
col_order <- c("company", "product_code", "product_category", "product_number", "address", "city", "country", "name")
refine <- refine[, col_order]
```

##(4) Add full address for geocoding
###(4b) Create a new column full_address that concatenates the 3 address fields (address, city, country), separated by columns
```{r}
library(tidyr)
refine <- unite(refine, address, city, country, col="full_address", sep=",")
```

##(5) Create dummy variables for company and product category
### (5b) Add four binary (1 or 0) columns for company: company_philips, company_akzo, company_van_houten and company_unilever
```{r}
install.packages("dplyr")
library(dplyr)
refine <- mutate(refine, company_phillips=refine$company[1:25])
View(refine)
refine$company_phillips[1:25] <- case_when(refine$company_phillips=="phillips" ~ 1, refine$company_phillips=="akzo" ~ 0, refine$company_phillips=="van houten" ~ 0, refine$company_phillips=="unilever" ~ 0)
View(refine)
refine <- mutate(refine, company_akzo=refine$company[1:25])
View(refine)
refine$company_akzo[1:25] <- case_when(refine$company_akzo=="phillips" ~ 0, refine$company_akzo=="akzo" ~ 1, refine$company_akzo=="van houten" ~ 0, refine$company_akzo=="unilever" ~ 0)
View(refine)
refine <- mutate(refine, company_van_houten=refine$company[1:25])
refine$company_van_houten[1:25] <- case_when(refine$company_van_houten=="phillips" ~ 0, refine$company_van_houten=="akzo" ~ 0, refine$company_van_houten=="van houten" ~ 1, refine$company_van_houten=="unilever" ~ 0)
View(refine)
refine <- mutate(refine, company_unilever=refine$company[1:25])
refine$company_unilever[1:25] <- case_when(refine$company_unilever=="phillips" ~ 0, refine$company_unilever=="akzo" ~ 0, refine$company_unilever=="van houten" ~ 0, refine$company_unilever=="unilever" ~ 1)
View(refine)
refine3 <- refine
View(refine3)
```
###(5c) Add four binary (1 or 0) columns for product category: product_smartphone, product_tv, product_laptop and product_tablet.
```{r}
new_cols_order <- c("company", "product_category", "product_code", "product_number", "product_smartphone","product_tv", "product_laptop", "product_tablet", "full_address", "name", "company_phillips", "company_akzo","company_van_houten","company_unilever")
refine <- refine[, new_cols_order]
refine <- mutate(refine, product_smartphone=refine$product_category[1:25])
View(refine)
refine$product_smartphone[1:25] <- case_when(refine$product_smartphone=="SmartPhone" ~ 1, refine$product_smartphone=="Laptop" ~ 0, refine$product_smartphone=="TV" ~ 0, refine$product_smartphone=="Tablet" ~ 0)
View(refine)
refine <- mutate(refine, product_laptop=refine$product_category[1:25])
refine$product_laptop[1:25] <- case_when(refine$product_laptop=="SmartPhone" ~ 0, refine$product_laptop=="Laptop" ~ 1, refine$product_laptop=="TV" ~ 0, refine$product_laptop=="Tablet" ~ 0)
View(refine)
refine <- mutate(refine, product_tv=refine$product_category[1:25])
refine$product_tv[1:25] <- case_when(refine$product_tv=="SmartPhone" ~ 0, refine$product_tv=="Laptop" ~ 0, refine$product_tv=="TV" ~ 1, refine$product_tv=="Tablet" ~ 0)
View(refine)
refine <- mutate(refine, product_tablet=refine$product_category[1:25])
refine$product_tablet[1:25] <- case_when(refine$product_tablet=="SmartPhone" ~ 0, refine$product_tablet=="Laptop" ~ 0, refine$product_tablet=="TV" ~ 0, refine$product_tablet=="Tablet" ~ 1)
```
## (6) Output refine_clean
```{r}
refine_clean2 <- refine
refine_clean2 <- write_csv(refine_clean2, "~/Documents/RProjects/Ch3_Ver2/refine_clean2.csv")
```








